---
title: "R Notebook"
output: html_notebook
---
# 0.Import library
```{r}
# The R library, cicero and monocle3 is NOT the part of celloracle package. Please install them yourself if you use this notebook.
# Please make sure that you installed cicero before running this script
rm(list=ls())
library(cicero)
library(monocle3)

```

# 1. Download data

This tutorial uses fetal brain scATAC-seq data from a 10x Genomics database. If you’re using your own scATAC-seq data, you will not need to download this dataset.
You can download the demo file with the following command. 

If the file download fails, please manually download and unzip the data.
http://cf.10xgenomics.com/samples/cell-atac/1.1.0/atac_v1_E18_brain_fresh_5k/atac_v1_E18_brain_fresh_5k_filtered_peak_bc_matrix.tar.gz

```{r}
# Create folder to store data
dir.create("data")

# Download demo dataset from 10x genomics 
download.file(url = "http://cf.10xgenomics.com/samples/cell-atac/1.1.0/atac_v1_E18_brain_fresh_5k/atac_v1_E18_brain_fresh_5k_filtered_peak_bc_matrix.tar.gz",
              destfile = "data/matrix.tar.gz")
# Unzip data
system("tar -xvf data/matrix.tar.gz -C data")
```

```{r}
# You can substitute the data path below to your scATAC-seq data.
data_folder <- "data/filtered_peak_bc_matrix"

# Create a folder to save results
output_folder <- "cicero_output"
dir.create(output_folder)
```


# 2. Load data and make Cell Data Set (CDS) object 
## 2.1. Process data to make CDS object

```{r}
# Read in matrix data using the Matrix package
indata <- Matrix::readMM(paste0(data_folder, "/matrix.mtx")) 
# Binarize the matrix
indata@x[indata@x > 0] <- 1

# Format cell info
cellinfo <- read.table(paste0(data_folder, "/barcodes.tsv"))
row.names(cellinfo) <- cellinfo$V1
names(cellinfo) <- "cells"

# Format peak info
peakinfo <- read.table(paste0(data_folder, "/peaks.bed"))
names(peakinfo) <- c("chr", "bp1", "bp2")
peakinfo$site_name <- paste(peakinfo$chr, peakinfo$bp1, peakinfo$bp2, sep="_")
row.names(peakinfo) <- peakinfo$site_name

row.names(indata) <- row.names(peakinfo)
colnames(indata) <- row.names(cellinfo)

# Make CDS
input_cds <-  suppressWarnings(new_cell_data_set(indata,
cell_metadata = cellinfo,
gene_metadata = peakinfo))

input_cds <- monocle3::detect_genes(input_cds)

#Ensure there are no peaks included with zero reads
input_cds <- input_cds[Matrix::rowSums(exprs(input_cds)) != 0,] 
```

# 3. Qauality check and Filtering
```{r}
# Visualize peak_count_per_cell
hist(Matrix::colSums(exprs(input_cds)))
```

```{r}
# Filter cells by peak_count
# Please set an appropriate threshold values according to your data 
max_count <-  15000
min_count <- 2000
input_cds <- input_cds[,Matrix::colSums(exprs(input_cds)) >= min_count] 
input_cds <- input_cds[,Matrix::colSums(exprs(input_cds)) <= max_count] 

```


# 4. Process cicero-CDS object

```{r}
# Data preprocessing
set.seed(2017)

input_cds <- detect_genes(input_cds)
input_cds <- estimate_size_factors(input_cds)
input_cds <- preprocess_cds(input_cds, method = "LSI")

# Dimensional reduction with umap
input_cds <- reduce_dimension(input_cds, reduction_method = 'UMAP', 
                              preprocess_method = "LSI")
umap_coords <- reducedDims(input_cds)$UMAP


cicero_cds <- make_cicero_cds(input_cds, reduced_coordinates = umap_coords)

# Save cds object if you want
#saveRDS(cicero_cds, paste0(output_folder, "/cicero_cds.Rds"))

```

# 5. Load reference genome information

To run cicero, you need to get a genomic coordinate file that contains the length of each chromosome.
You can download the mm10 genomic information with the following command.

If your scATAC-seq data was aligned to a different reference genome, you will need the reference-specific genome coordinates. See the Cicero documentation for more information.

https://cole-trapnell-lab.github.io/cicero-release/docs_m3/#installing-cicero

```{r}
# !!Please make sure that the reference genome information below matches the reference genome of your scATAC-seq data.

# If your scATAC-seq uses mm10 reference genome, you can read chromosome length file with the following command.
download.file(url = "https://raw.githubusercontent.com/morris-lab/CellOracle/master/docs/demo_data/mm10_chromosome_length.txt",
              destfile = "./mm10_chromosome_length.txt")
chromosome_length <- read.table("./mm10_chromosome_length.txt")

# For mm9 genome, you can use the following command.
#data("mouse.mm9.genome")
#chromosome_length <- mouse.mm9.genome

# For hg19 genome, you can use the following command.
#data("human.hg19.genome")
#chromosome_length <- mhuman.hg19.genome


```

# 6. Run Cicero

```{r}
# Run the main function
conns <- run_cicero(cicero_cds, chromosome_length) # Takes a few minutes to run

# Save results if you want
#saveRDS(conns, paste0(output_folder, "/cicero_connections.Rds"))

# Check results
head(conns)
```

# 7. Save results for the next step

```{r}
all_peaks <- row.names(exprs(input_cds))
write.csv(x = all_peaks, file = paste0(output_folder, "/all_peaks.csv"))
write.csv(x = conns, file = paste0(output_folder, "/cicero_connections.csv"))

```


